The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-10 rational.tex
HOF

\newcommand{\lcal}{\ensuremath{\lambda}-calculus}
\newcommand{\G}{\ensuremath{\mathcal{G}}\xspace}

@@

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! langRev = \Pi
%! langRevT = \Pi^{o}
%! langRevEE = \Pi^{\eta\epsilon}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Computing in the Field of Algebraic Numbers}

@@

\item Do we really have HOF and continuations? What is the exact
  relationship? Do we have linearly used functions? Do we have linear
  functions (in that they use their argument linearly)?

\item Can we complete the construction in \cite{infeffects} to be full
  STLC? This is again a HOF question in another guise.

\item What are information effects in {{langRevEE}}? Depending on what
  we think about HOF, do we have a monad or is it still an arrow?

\subsection{Higher Order Functions}

These functions aren't really values. There is no value one can
produce that denotes a function. These functions are the ability to
transform a value - the possibility of transforming a value.  In the
product encoding of environments, there is no value one can assign to
a variable such that it denotes a function.

Actually this is possible. Consider two functions {{f : b1 <-> b2}}
and {{g : b1 <-> b2}}. A value of type {{bool}} is sufficient to
discriminate them. Hence the {{boo}} is the first class representation
of the functions and can be thought of as the address of the function.

@@

\subsection{Other}
To summarize negative, fractional, square root, and imaginary types all make
sense. What they help you accomplish as a programmer is to disassociate
global invariants into local ones that can be satisfied independently by
subcomputations with no synchronization or communication. A computation
producing something of type $a/b$ does not need to concern itself with who is
going to supply the missing $b$: it just does its part. Conversely faced with
a complicated task, a computation might decide to break it into pieces and
demand these pieces using negative types.

>>>>

\newcommand{\lcal}{\ensuremath{\lambda}-calculus\xspace}
\newcommand{\G}{\ensuremath{\mathcal{G}}\xspace}

@@

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{*}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o+ = \multimap^{+}
%! -o* = \multimap^{\times}
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! langRev = \Pi
%! langRevT = \Pi^{o}
%! langRevEE = \Pi^{\eta\epsilon}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Conventional Higher Order Functions}

In a conventional language, higher-order functions capture scope and
can be represented as first class values. On the other hand,
{{langRevEE}} gives us two notions of functions both of which share
similarities with conventional HOFs.

\begin{itemize}
\item 
Additive functions are not really values, but are possible paths of
execution. However, they can be encoded as values as follows: Consider
two functions {{f:b1 -o+ b2}} and {{g:b1-o+b2}}. A value of type
{{bool}} can discriminate them and hence a value of type {{bool}} acts
as the `address' of the function and is a first class value
representing one of the {{b1 -o+ b2}} functions.  This is similar to
how first classes functions a'la \lcal are represented by their
addresses when compiled to conventional hardware.

\item
Multiplicative functions, of the form {{b1 -o* b2}}, on the other hand
are indeed first class values. However these functions can only be
linearly used -- i.e. they represent a contraint that can be satisfied
at most once in one `computational world'. It may be possible that to
repeatedly invoke multiplicative functions with an appropriate
recursive type whose unfoldings allow for repeated unification.
\end{itemize}

The exact encoding on HOF in {{langRevEE}} should shed light on
several problems, including the connections to continuations discussed
in Sec. \ref{sec:related}. They should also allow for a natural
encoding of HOFs in the translations of \lcal to {{langRevT}} and
indicate a monadic encapsulation of of information effects which are
currrently structured as an arrow.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing in the Field of Algebraic Numbers}

@@

@@

% \subsection{Other}

To summarize negative, fractional, square root, and imaginary types all make
sense. What they help you accomplish as a programmer is to disassociate
global invariants into local ones that can be satisfied independently by
subcomputations with no synchronization or communication. A computation
producing something of type $a/b$ does not need to concern itself with who is
going to supply the missing $b$: it just does its part. Conversely faced with
a complicated task, a computation might decide to break it into pieces and
demand these pieces using negative types.
